---
title: "Relational Data Assignment"
author: "Adira Dangoy"
date: "November 9, 2017"
output: html_document
---
Mutating joins, which add new variables to one data frame from matching observations in another.

Filtering joins, which filter observations from one data frame based on whether or not they match an observation in the other table.

Set operations, which treat observations as if they were set elements.

```{r}
library(tidyverse)
library(nycflights13)
```
13.2.1 Exercises
1.) Variables: tail number, origin, dest, carrier.
Tables: arilines, planes, airports.

2.) Weather and airports are connected with origin. Origin on weather should have an arrow pointing to origin on flights which is a sub for airports. 

3.)It would have the additional relation with dest in flights.

4.) Primary keys would be month and day from the weather table. I would create these into a new table only for special days.

_________________
A primary key uniquely identifies an observation in its own table. For example, planes$tailnum is a primary key because it uniquely identifies each plane in the planes table.

A foreign key uniquely identifies an observation in another table. For example, the flights$tailnum is a foreign key because it appears in the flights table where it matches each flight to a unique plane.

13.3.1 Exercises
1.) 
```{r}
flights %>% 
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_id = row_number()) 
```

2.)
```{r}
Lahman::Batting
```

Primary Keys: YearID, stint, and TeamID

```{r}
babynames::babynames
```
Primary Keys: year, sex, names

```{r}
nasaweather::atmos
```
Primary Keys: lat, long, year, month.

```{r}
fueleconomy::vehicles
```
Primary Key: ID
```{r}
ggplot2::diamonds
```
No primary key.




3.)
```{r}
Lahman::Batting
```
Batting
primary key: playerID, yearID, stint
foreign keys:
playerID -> Master.playerID

```{r}
Lahman::Master
```
Master
Primary Key: playerID

```{r}
Lahman::Salaries
```
Salaries
Primary Key: yearID, teamID, PlayerID
Foreign key: playerID>master.playerID.

```{r}
Lahman::Managers
```
Managers
Primary Key: playerID,yearID,teamID.
Foreign Key:playerID > Master.teamID

```{r}
Lahman::AwardsManagers
```

Awards Managers
Primary Key: playerID, awardID, yearID
playerID -> Master.playerID
playerID, yearID, lgID > Managers.playerID, yearID, lgID

```{r}
Lahman::Batting
```

```{r}
Lahman::Pitching
```

```{r}
Lahman::Fielding
```
All three tables have PlayerIDs, YearIDs, Stints, and TeamIDs. So they can all be traces back to eachother and you can compare the player IDs to see their stats for all three catagories.

__________

```{r}
nycflights13::flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(carrier, sort = TRUE)
```
```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```
1.)
```{r}
avg_dest_delays <-
  flights %>%
  group_by(dest) %>%
  
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c(dest = "faa"))

avg_dest_delays %>%
  ggplot(aes(lon, lat, colour = delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

 2.)
 

```{R}
flights %>%
  left_join(airports, by = c(dest = "faa")) %>%
  left_join(airports, by = c(origin = "faa")) %>%
  head()
```


3.)

```{r}
plane_ages <- 
  planes %>%
  mutate(age = 2013 - year) %>%
  select(tailnum, age)

flights %>%
  inner_join(plane_ages, by = "tailnum") %>%
  group_by(age) %>%
  filter(!is.na(dep_delay)) %>%
  summarise(delay = mean(dep_delay)) %>%
  ggplot(aes(x = age, y = delay)) +
  geom_point() +
  geom_line()
```
```{r}
flight_weather <-
  flights %>%
  inner_join(weather, by = c("origin" = "origin",
                            "year" = "year",
                            "month" = "month",
                            "day" = "day",
                            "hour" = "hour"))

flight_weather %>%
  group_by(precip) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = precip, y = delay)) +
    geom_line() + geom_point()
```


5.)
```{r}
library(viridis)
```

```{r}
flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap() + 
  scale_color_viridis()
```

There was a large series of derechos that caused several tornados in the US you can see several delays in the south where it touched and one large point in the south during the storms had 150 delays.
___________


1.)
```{r}
flights %>%
  anti_join(planes, by = "tailnum") %>%
count
```
```{r}
planes
```
AA and MQ don't Use tail numbers

2.)
```{r}
planes_100Flights <- 
  filter(flights) %>%
  group_by(tailnum) %>%
  count() %>%
  filter(n > 100)

flights %>%
  semi_join(planes_gt100, by = "tailnum")
```

3.)
```{r}
(fueleconomy::vehicles)
```
```{r}
(fueleconomy::common)
```

```{r}
fueleconomy::vehicles %>%
  semi_join(fueleconomy::common, by = c("make", "model"))
```
4.)
  
```{r}
anti_join(flights, airports, by = c("dest" = "faa"))
```
This shows the flights that go to an airport that is not in FAA list of destinations.
```{r}
anti_join(airports, flights, by = c("faa" = "dest"))
```

These are US airports that don’t have a flight in the data. This shows us that there were no flights to that aiport from New York during this period.

5.)I believe there is a unique relationship between the two. All tailnumbers have a unique pattern and each airline has their own planes. There is a chance that the carriers could merge or sell planes to other carriers but this isn't really shown in our data.
```{r}
flights %>%
  group_by(tailnum, carrier) %>%
  count() %>%
  filter(n() > 1) %>%
  select(tailnum) %>%
  distinct()
```


