---
title: "Relational Data Assignment"
author: "Adira Dangoy"
date: "November 9, 2017"
output: html_document
---
Mutating joins, which add new variables to one data frame from matching observations in another.

Filtering joins, which filter observations from one data frame based on whether or not they match an observation in the other table.

Set operations, which treat observations as if they were set elements.

```{r}
library(tidyverse)
library(nycflights13)
```
13.2.1 Exercises
1.) Variables: tail number, origin, dest, carrier.
Tables: arilines, planes, airports.

2.) Weather and airports are connected with origin. Origin on weather should have an arrow pointing to origin on flights which is a sub for airports. 

3.)It would have the additional relation with dest in flights.

4.) Primary keys would be month and day from the weather table. I would create these into a new table only for special days.

_________________
A primary key uniquely identifies an observation in its own table. For example, planes$tailnum is a primary key because it uniquely identifies each plane in the planes table.

A foreign key uniquely identifies an observation in another table. For example, the flights$tailnum is a foreign key because it appears in the flights table where it matches each flight to a unique plane.

13.3.1 Exercises
1.) 
```{r}
flights %>% 
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_id = row_number()) 
```

2.)
```{r}
Lahman::Batting
```

Primary Keys: YearID, stint, and TeamID

```{r}
babynames::babynames
```
Primary Keys: year, sex, names

```{r}
nasaweather::atmos
```
Primary Keys: lat, long, year, month.

```{r}
fueleconomy::vehicles
```
Primary Key: ID
```{r}
ggplot2::diamonds
```
No primary key.




3.)
```{r}
Lahman::Batting
```
```{r}
Lahman::Master
```
```{r}
Lahman::Salaries
```


